---
title: "Discrepancy exploration"
subtitle: "R code"
author: "Arnald Puy"
header-includes:
  - \usepackage[font=footnotesize]{caption}
  - \usepackage{dirtytalk}
  - \usepackage{booktabs}
  - \usepackage{tabulary}
  - \usepackage{enumitem}
  - \usepackage{lmodern}
  - \usepackage[T1]{fontenc}
  - \usepackage{tikz}
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
    keep_tex: true
  word_document:
    toc: no
    toc_depth: '2'
  html_document:
    keep_md: true
link-citations: yes
fontsize: 11pt
bibliography: /Users/arnaldpuy/Documents/bibtex/ERC-sensitivity.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = "tikz", cache = TRUE)
```

\newpage

# Preliminary functions

```{r, results="hide", message=FALSE, warning=FALSE, cache=FALSE}

# PRELIMINARY ------------------------------------------------------------------

# Function to read in all required packages in one go:
loadPackages <- function(x) {
  for(i in x) {
    if(!require(i, character.only = TRUE)) {
      install.packages(i, dependencies = TRUE)
      library(i, character.only = TRUE)
    }
  }
}

theme_AP <- function() {
  theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          legend.background = element_rect(fill = "transparent",
                                           color = NA),
          legend.margin=margin(0, 0, 0, 0),
          legend.box.margin=margin(-7,-7,-7,-7), 
          legend.key = element_rect(fill = "transparent",
                                    color = NA), 
          strip.background = element_rect(fill = "white"))
}

# Load the packages
loadPackages(c("sensobol", "data.table", "tidyverse", "parallel", 
               "RcppAlgos", "scales", "doParallel", "benchmarkme"))
```

\newpage

# Goal
We aim at checking whether discrepancy measures can be used as a sensitivity measure to reliably rank parameters in terms of their influence in the model output. To that aim, we explore how well the symmetric L2-discrepancy measure ranks the most important parameters of a set of functions, including for the moment the Oakley and O'Hagan 2004 and the Bratley et al. 1992 functions, and compare its efficacy with that of the Jansen estimator. We do the comparison on Savage-score transformed ranks [@Iman1987].

The discrepancy is computed directly on the plane of the $X_i$, $Y$ scatterplots where $X_i$ is a coordinate $(0,1)$ of the input hypercube and $Y$ is the output rescaled in $0,1$. For each of the $k$ scatterplots, the points are sorted by one of the  $X_i$. The idea is that a ''pattern'' (=effect) corresponds to a deviation from a uniform distribution of points (noise) for a non-influential variable: the largest the effect, the larger the discrepancy, e.g., there are areas in the plot where points are denser and areas where they are more rarefied.  

One might wonder why use a discrepancy measure when there are the Sobol' indices. The reply is that many do not understand the theory behind them (and the same applies to moment independent methods or VARS), while everyone understands scatterplots. 'Explaining' discrepancy as a measure of the non-uniformity of the scatterplots appears an avenue likely to appeal to some users.

# Required functions

## Savage ranks function

```{r savage_fun}

# SAVAGE SCORES FUNCTION -------------------------------------------------------

savage_scores <- function(x) {
  true.ranks <- rank(-x)
  p <- sort(1 / true.ranks)
  mat <- matrix(rep(p, length(p)), nrow = length(p), byrow = TRUE)
  mat[upper.tri(mat)] <- 0
  out <- sort(rowSums(mat), decreasing = TRUE)[true.ranks]
  return(out)
}
```

## Discrepancy function

```{r discrepancy_fun, dependson="savage_scores"}

# DISCREPANCY FUNCTION ---------------------------------------------------------

# This is based on the function discrepancyCriteria_cpp of
# the sensitivity package

# Function to rescale ---------------------------
rescale_fun <- function(x) (x - min(x)) / (max(x) - min(x))

# Source cpp code -------------------------------
Rcpp::sourceCpp("cpp_functions.cpp")

# Function --------------------------------------
discrepancy_s2 <- function (design) {

  X <- as.matrix(design)
  dimension <- ncol(X)
  n <- nrow(X)
  
  # Reescale if needed-------
  
  if (min(X) < 0 || max(X) > 1) {
    
    X <- apply(X, 2, rescale_fun)
  }
  
  # Compute discrepancy
  P <- 1 + 2 * X - 2 * X^2
  s1 <- DisS2_Rowprod(t(P), dimension)
  s2 <- DisS2_Crossprod(c(t(X)), dimension)
  R <- sqrt(((4/3)^dimension) - ((2/n) * s1) + ((2^dimension/n^2) * s2))
  return(R)
}

# Discrepancy function --------------------------
discrepancy <- function(mat, y, params) {
  value <- sapply(1:ncol(mat), function(j) {
    design <- cbind(mat[, j], y) 
    value <- discrepancy_s2(design = design)
  })
  return(value)
}
```

## Jansen estimator

```{r jansen_fun}

# FUNCTION TO COMPUTE JANSEN TI ------------------------------------------------

jansen_ti <- function(d, N, params) {
  m <- matrix(d, nrow = N)
  k <- length(params)
  Y_A <- m[, 1]
  Y_AB <- m[, -1]
  f0 <- (1 / length(Y_A)) * sum(Y_A)
  VY <- 1 / length(Y_A) * sum((Y_A - f0) ^ 2)
  value <- (1 / (2 * N) * Rfast::colsums((Y_A - Y_AB) ^ 2)) / VY
  return(value)
}
```

## Function to replicate sample matrix

```{r replicas}

# FUNCTION TO CREATE REPLICAS OF SAMPLE MATRIX ---------------------------------

# For discrepancy
CutBySize <- function(m, block.size, nb = ceiling(m / block.size)) {
  int <- m / nb
  upper <- round(1:nb * int)
  lower <- c(1, upper[-nb] + 1)
  size <- c(upper[1], diff(upper))
  cbind(lower, upper, size)
}

sobol_replicas <- function(matrices, N, params, replicas,...) {
  mat <- sobol_matrices(matrices = matrices, N = N * replicas, 
                        params = params,...)
  index <- CutBySize(nrow(mat), block.size = N)
  out <- list()
  for (i in 1:nrow(index)) {
    out[[i]] <- mat[index[i, "lower"]:index[i, "upper"], ]
  }
  return(out)
}

# For Sobol' indices
scrambled_sobol <- function(A, B) {
  X <- rbind(A, B)
  for(i in 1:ncol(A)) {
    AB <- A
    AB[, i] <- B[, i]
    X <- rbind(X, AB)
  }
  AB <- rbind(A, X[((2*nrow(A)) + 1):nrow(X), ])
  return(AB)
}
# Function to create replicas of the A, B and AB matrices
sobol_replicas_A_AB <- function(N, params, replicas) {
  k <- length(params)
  df <- randtoolbox::sobol(n = N * replicas, dim = k * 2)
  indices <- CutBySize(nrow(df), nb = replicas)
  X <- A <- B <- out <- list()
  for(i in 1:nrow(indices)) {
    lower <- indices[i, "lower"]
    upper <- indices[i, "upper"]
    X[[i]] <- df[lower:upper, ]
  }
  for(i in seq_along(X)) {
    A[[i]] <- X[[i]][, 1:k]
    B[[i]] <- X[[i]][, (k + 1) : (k * 2)]
  }
  for(i in seq_along(A)) {
    out[[i]] <- scrambled_sobol(A[[i]], B[[i]])
  }
  return(out)
}
```

## Function to check correlation between discrepancy and jansen

```{r correlation, dependson=c("replicas", "jansen_fun", "discrepancy_fun", "savage_fun")}

# FUNCTION TO CHECK CORRELAITON BETWEEN DISCREPANCIES AND SAVAGE SCORES --------

cor_fun <- function(N = N, params = params, replicas = replicas, approach = NULL,
                    model_fun = NULL, true_scores = NULL,...) {
  
  if (is.null(approach)) {
    stop("approach should be either discrepancy or sobol")
  }
  
  if (is.null(model_fun) | is.null(true_scores)) {
    stop("A function and the true ranks should be provided")
  }
  
  if (approach == "discrepancy") {
    
    mat <- sobol_replicas(matrices = "A", N = N, params = params, 
                          replicas = replicas,...)
    y <- lapply(mat, model_fun)
    
    out <- list()
    for(i in 1:length(y)) {
      out[[i]] <- discrepancy(mat = mat[[i]], y = y[[i]], params = params) %>%
        savage_scores()
    }
    
  } else if (approach == "sobol") {
    
    mat <- sobol_replicas_A_AB(N = N, params = params, replicas = replicas)
    y <- lapply(mat, model_fun)
    out <- lapply(y, function(x) jansen_ti(d = x, N = N, params = params) %>%
                    savage_scores()) 
  }
  
  final <- unlist(lapply(out, function(x) cor(x, true_scores)))
  
  return(final)
}
```

# The Bratley et al. 1992 and the Oakland and O'Hagan 2004 functions

```{r bratley_check}

# CHECK SOBOL' INDICES AND SCATTERPLOT FOR SEVERAL FUNCTIONS ------------------

N <- 2^14 # Base sample size
matrices <- c("A", "B", "AB")
R <- 10^3 # Bootstrap replicas

models <- c("Bratley et al. 1992", "Oakland and O'Hagan 2004")
fun_list <- list(bratley1992_Fun, oakley_Fun)
names(fun_list) <- models

out <- ind <- savage.rank <- list()
for (i in models) {
  
  if (i == "Bratley et al. 1992") {
    params <- paste("X", 1:6, sep = "")
    
  } else if (i == "Oakland and O'Hagan 2004") {
    params <- paste("X", 1:15, sep = "")
  }
  
  mat <- sobol_matrices(matrices = matrices, N = N, params = params)
  y <- fun_list[[i]](mat)
  ind[[i]] <- sobol_indices(matrices = matrices, Y = y, N = N, params = params, 
                       boot = TRUE, R = R)
  savage.rank[[i]] <- ind[[i]]$results[sensitivity == "Ti"] %>%
    .[, savage_scores(original)]
  out[[i]] <-plot_scatter(data = mat, N = N, Y = y, params = params) 
}

lapply(ind, plot)
```

# Run the model

```{r model, dependson="correlation"}

# THE MODEL -------------------------------------------------------------------

matrices <- "A"
replicas <- 150 # Number of replicas of the sample matrix

output <-  list()

for(i in models) {
  
  if (i == "Bratley et al. 1992") {
    params <- paste("X", 1:6, sep = "")
    
  } else if (i == "Oakland and O'Hagan 2004") {
    params <- paste("X", 1:15, sep = "")

  }
  
  for (j in c("discrepancy", "sobol")) {
    
    if (j == "discrepancy") {
      
      if(i == "Bratley et al. 1992") {
        sample.sizes <- seq(4, 150, by = 4)
        final.sample.sizes <- sample.sizes
      } else {
        sample.sizes <- seq(4, 320, by = 4)
        final.sample.sizes <- sample.sizes
      }

    } else if (j == "sobol") {
      sample.sizes <- 2:20
      final.sample.sizes <- sample.sizes * (length(params) + 1)
      
    }
    output[[i]][[j]] <- mclapply(sample.sizes, function(x) 
      cor_fun(N = x, params = params, replicas = replicas, model_fun = fun_list[[i]], 
              true_scores = savage.rank[[i]], approach = j, 
              scrambling = 1), 
      mc.cores = detectCores() * 0.75)
    
    names(output[[i]][[j]]) <- final.sample.sizes

  }
}

# Arrange output -------------------
results <- lapply(output, function(x) lapply(x, function(y) 
  lapply(y, function(z) data.table(z))))  %>%
  lapply(., function(x) lapply(x, function(y) rbindlist(y, idcol = "Model.runs"))) %>%
  lapply(., function(x) rbindlist(x, idcol = "Approach")) %>%
  rbindlist(., idcol = "Function") %>%
  .[, Model.runs:= as.numeric(Model.runs)] %>%
  .[, Approach:= ifelse(Approach == "discrepancy", "Discrepancy", "Jansen estimator")] %>%
  na.omit()

```

```{r plot, dependson="model", warning=FALSE, fig.height=4, fig.cap="Correlation between the Savage scores obtained with $T_i$ (Jansen estimator) and those obtained with discrepancy."}

# Plot the results ------------------------------------------------------------

# New facet label names for supp variable
supp.labs <- c("Symmetric $L2$ discrepancy", "Jansen estimator")
names(supp.labs) <- c("Discrepancy", "Jansen estimator")

ggplot(results, aes(Model.runs, z, group = Model.runs)) +
  geom_boxplot(position = "identity", outlier.size = 0.5) +
  labs(x = "Nº of model runs", y = "Correlation") +
  facet_grid(Approach~Function, 
             scale = "free_x", space = "free", 
             labeller = labeller(Approach = supp.labs)) +
  theme_AP()
```

# The metamodel

```{r metamodel_fun}

# RANDOM FUNCTIONS AND DISTRIBUTIONS ------------------------------------------

# Functions ------------------------------------
function_list <- list(
  Linear = function(x) x,
  Quadratic = function(x) x ^ 2,
  Cubic = function(x) x ^ 3,
  Exponential = function(x) exp(1) ^ x / (exp(1) - 1),
  Periodic = function(x) sin(2 * pi * x) / 2,
  Discontinuous = function(x) ifelse(x > 0.5, 1, 0),
  Non.monotonic = function(x) 4 * (x - 0.5) ^ 2,
  Inverse = function(x) (10 - 1 / 1.1) ^ -1 * (x + 0.1) ^ - 1, 
  No.effect = function(x) x * 0, 
  Trigonometric = function(x) cos(x), 
  Piecewise.large = function(x) ((-1) ^ as.integer(4 * x) * 
                                   (0.125 - (x %% 0.25)) + 0.125),
  Piecewise.small = function(x) ((-1) ^ as.integer(32 * x) * 
                                   (0.03125 - 2 * (x %% 0.03125)) + 0.03125) / 2,
  Oscillation = function(x) x ^ 2 - 0.2 * cos(7 * pi * x)
)

# Random distributions  -------------------------
sample_distributions <- list(
  "uniform" = function(x) x,
  "normal" = function(x) qnorm(x, 0.5, 0.15),
  "beta" = function(x) qbeta(x, 8, 2),
  "beta2" = function(x) qbeta(x, 2, 8),
  "beta3" = function(x) qbeta(x, 2, 0.8),
  "beta4" = function(x) qbeta(x, 0.8, 2),
  "logitnormal" = function(x) logitnorm::qlogitnorm(x, 0, 3.16)
  # Logit-normal, Bates too?
)

random_distributions <- function(X, phi) {
  names_ff <- names(sample_distributions)
  if(!phi == length(names_ff) + 1) {
    out <- sample_distributions[[names_ff[phi]]](X)
  } else {
    temp <- sample(names_ff, ncol(X), replace = TRUE)
    out <- sapply(seq_along(temp), function(x) 
      sample_distributions[[temp[x]]](X[, x]))
  }
  return(out)
}
```

```{r model_fun, dependson=c("metamodel_fun", "savage_fun", "discrepancy_fun", "jansen_fun")}

# DEFINE MODEL ----------------------------------------------------------------

model_fun <- function(tau, epsilon, base.sample.size, cost.discrepancy, phi, k) {
  params <- paste("X", 1:k, sep = "")
  
  if (tau == 1) {
    
    type <- "R"
    
  } else if (tau == 2) {
    
    type <- "QRN"
  }
  set.seed(epsilon)
  discrepancy.mat <- sobol_matrices(matrices = "A", N = cost.discrepancy, 
                                    params = params, 
                                    type = type)
  
  set.seed(epsilon)
  jansen.mat <- sobol_matrices(matrices = c("A", "AB"), N = base.sample.size, 
                               params = params, 
                               type = type)
  
  set.seed(epsilon)
  #all.matrices <- random_distributions(X = rbind(discrepancy.mat, jansen.mat), phi = phi)
  
  all.matrices <- rbind(discrepancy.mat, jansen.mat)
  
  y <- sensobol::metafunction(data = all.matrices, epsilon = epsilon)
  
  discrepancy.value <- discrepancy(mat = all.matrices[1:cost.discrepancy, ], 
                                   y = y[1:cost.discrepancy], params = params)
  
  jansen.value <- jansen_ti(d = y[(cost.discrepancy + 1):length(y)],
                            N = base.sample.size, params = params)
  
  savage.discrepancy <- savage_scores(discrepancy.value)
  jansen.discrepancy <- savage_scores(jansen.value)
  
  out <- cor(savage.discrepancy, jansen.discrepancy)
  
  return(out)
}
```

```{r matrix}

# CREATE SAMPLE MATRIX -------------------------------------------------------

N <- 2^10
params <- c("epsilon", "phi", "k", "tau", "base.sample.size")
mat <- sobol_matrices(matrices = "A", N = N, params = params)

# Define distributions --------------------------

mat[, "epsilon"] <- floor(qunif(mat[, "epsilon"], 1, 200))
mat[, "phi"] <- floor(mat[, "phi"] * 8) + 1
mat[, "k"] <- floor(qunif(mat[, "k"], 3, 50))
mat[, "tau"] <- floor(mat[, "tau"] * 2) + 1
mat[, "base.sample.size"] <- floor(qunif(mat[, "base.sample.size"], 10, 100))

cost.jansen <- mat[, "base.sample.size"] * (mat[, "k"] + 1)
cost.discrepancy <- cost.jansen

final.mat <- cbind(mat, cost.jansen, cost.discrepancy)
```

```{r run_model_fun, dependson="model_fun"}

# RUN MODEL  ------------------------------------------------------------------

y <- mclapply(1:nrow(final.mat), function(i) {
  model_fun(tau = final.mat[i, "tau"],
            epsilon = final.mat[i, "epsilon"], 
            base.sample.size = final.mat[i, "base.sample.size"], 
            cost.discrepancy = final.mat[i, "cost.discrepancy"], 
            phi = final.mat[i, "phi"], 
            k = final.mat[i, "k"])}, 
  mc.cores = floor(detectCores() * 0.75))

y <- unlist(y)
output <- data.table(cbind(final.mat, y))
```

```{r plot_uncertainty, dependson="run_model_fun", warning = FALSE, fig.height=3, fig.width=3}

# PLOT UNCERTAINTY ------------------------------------------------------------

ggplot(output, aes(cost.discrepancy, k, color = y)) + 
  geom_point() + 
  scale_colour_gradientn(colours = c("red", "orange", "yellow", "lightgreen"), 
                         name = expression(italic(r)), 
                         breaks = pretty_breaks(n = 3)) +
  scale_x_continuous(breaks = pretty_breaks(n = 3)) +
  scale_y_continuous(breaks = pretty_breaks(n = 3)) +
  labs(x = "Nº of model runs", 
       y = "$k$") + 
  theme_AP() + 
  theme(legend.position = "top", 
        strip.background = element_rect(fill = "white"))
```

\newpage

# Session information

```{r session_information}

# SESSION INFORMATION --------------------------------------------------------------

sessionInfo()

## Return the machine CPU
cat("Machine:     "); print(get_cpu()$model_name)

## Return number of true cores
cat("Num cores:   "); print(detectCores(logical = FALSE))

## Return number of threads
cat("Num threads: "); print(detectCores(logical = TRUE))

## Return the machine RAM
cat("RAM:         "); print (get_ram()); cat("\n")
```

\newpage

# References